Big Ideas in Sports Analytics and Statistical Tools for their Investigation

### 2022-07-11 sports_analytics.Rmd
some notes on discrete sports

@Pattani2012 describes ESPN's expected points metric.

@@

@Yurko2019, which uses an expected points model as part of the nfl war framework, uses down (1-4, seconds remaining in half, yards from endzone, yards to go (log), indicator for goal to go (GTG), and indicator for under 2 minutes (UTM).

@@

Even in sports where the concept of a state is not so easy to define, the value of a possession can be estimated, with the help of tracking data. Over the past decade or so, professional sports leagues have collected tracking data which record the locations of all players and the ball (or puck) throughout a game. This sophisticated type of data allows researchers to produce advanced analyses of the captured spatio-temporal information and better understand the game, which is much more difficult to be accomplished with pre-existing resources such as box-score results and play-by-play data.

In basketball, @cervone2014pointwise and @cervone2016multiresolution introduce expected possession value (EPV) as an assessment of a player's on-court performance. This metric is a continuous-time estimate of the expected number of points for the offensive team on a given possession using player and ball locations. The EPV takes into account all possible outcomes (a shot attempt, a pass, etc.) for a given player with the ball, with different weights being assigned to each decision.

>>>>

@@

@Yurko2019, which uses an expected points model as part of the nfl war framework, uses down (1-4, seconds remaining in half, yards from endzone, yards to go (log), indicator for goal to go (GTG), and indicator for under 2 minutes (UTM)

multinomial logistic regression
next score r.v.

@@

Even in sports where the concept of a state is not so easy to define, the value of a possession can be estimated, with the help of tracking data. 
Over the past decade or so, professional sports leagues have collected tracking data which record the locations of all players and the ball (or puck) throughout a game. 
This sophisticated type of data allows researchers to produce advanced analyses of the captured spatio-temporal information and better understand the game, which is much more difficult to be accomplished with pre-existing resources such as traditional box-score results and play-by-play data.

In basketball, @cervone2014pointwise and @cervone2016multiresolution introduce expected possession value (EPV) as an assessment of a player's on-court performance. 
This metric is a continuous-time estimate of the expected number of points for the offensive team on a given possession using player and ball locations. 
The EPV takes into account all possible outcomes (a shot attempt, a pass, etc.) for a given player with the ball, with different weights being assigned to each decision. 
Consequently, the authors derive a metric called EPV-Added (EPVA), measuring a player's EPV contribution in a given situation relative to a league-average player.

(include code demo? https://github.com/dcervone/EPVDemo)
